---
title: "The Methylome in Female Adolescent Conduct Disorder: Neural Pathomechanisms and Environmental Risk Factors"
subtitle: 'Post-hoc Analyses'
author: "AG Chiocchetti"
date: "29 Dezember 2020"
output: html_document
code_folding: hide
---

```{r setup, echo=F, include=FALSE}
library(RColorBrewer)
library(corrplot)
library(gplots)
library(ggplot2)
library(lavaan)
library(semPlot)
library(tidyverse)
library(DESeq2)
library(GenomicRanges)
library(knitr)
library(glmpca)
library(kableExtra)
library(BiocParallel)
library(gprofiler2)
library(xlsx)
library(org.Hs.eg.db)
library(visNetwork)
library(webshot); #webshot::install_phantomjs() #in case phantomjs was not installed 

Home = getwd()

thresholdp = 10e-3
thresholdLFC = 0.5


source(paste0(Home, "/code/custom_functions.R"))

options(stringsAsFactors = F)
opts_chunk$set(echo = TRUE, 
               fig.height = 7, 
               fig.width = 9, 
               message = TRUE,
               warning = TRUE,
               fig.align='center',
               dev = c("png", "svg"))


Dark8 = brewer.pal(8, "Dark2")
Dark8_50 = paste0(brewer.pal(8, "Dark2"), "7D")
jetcolors = colorRampPalette(c("darkblue", "skyblue", "green", "yellow", "orange", "red", "darkred"))
OBcolors = colorRampPalette(c("darkblue", "skyblue", "white",  "orange", "darkorange3"))
load(paste0(Home,"/output/dds_filt_analyzed.RData"))
load(paste0(Home,"/output/resdmr.RData"))
load(paste0(Home,"/output/resultsdmr_table.RData"))

results_Deseq = results(dds_filt)
log2_cpm = log2(counts(dds_filt, normalized=T)+1)



```

## Genomic location enrichment

Significant loci with a p-value <= `r thresholdp` and a absolute log2 fold-change lager `r thresholdLFC` were tested for enrichment in annotated genomic feature using fisher exact test.  


```{r genomiclocation, fig.width=15, fig.height=5}

Ranges=rowData(dds_filt)

TotTagsofInterest=sum(Ranges$WaldPvalue_groupCD<=thresholdp & abs(Ranges$groupCD)>thresholdLFC)

Resall=data.frame()
index = Ranges$WaldPvalue_groupCD<=thresholdp& abs(Ranges$groupCD)>thresholdLFC
for (feat in unique(Ranges$feature)){
  tmp=table(Ranges$feature == feat, signif=index)
  resfish=fisher.test(tmp)
  res = c(resfish$estimate, unlist(resfish$conf.int), resfish$p.value)
  Resall = rbind(Resall, res)
}
tmp=table(Ranges$tf_binding!="", signif=index)
resfish=fisher.test(tmp)
res = c(resfish$estimate, unlist(resfish$conf.int), resfish$p.value)
Resall = rbind(Resall, res)
tmp=table(Ranges$cpg=="cpg", signif=index)
resfish=fisher.test(tmp)
res = c(resfish$estimate, unlist(resfish$conf.int), resfish$p.value)
Resall = rbind(Resall, res)
colnames(Resall)=c("OR", "CI95L", "CI95U", "P")
rownames(Resall)=c(unique(Ranges$feature), "TF-binding", "CpG-island")
Resall$Beta = log(Resall$OR)
Resall$SE = (log(Resall$OR)-log(Resall$CI95L))/1.96
Resall$Padj=p.adjust(Resall$P, method = "bonferroni")

Resdown=data.frame()
index = Ranges$WaldPvalue_groupCD<=thresholdp & Ranges$groupCD<thresholdLFC
for (feat in unique(Ranges$feature)){
  tmp=table(Ranges$feature == feat, signif=index)
  resfish=fisher.test(tmp)
  res = c(resfish$estimate, unlist(resfish$conf.int), resfish$p.value)
  Resdown = rbind(Resdown, res)
}
tmp=table(Ranges$tf_binding!="", signif=index)
resfish=fisher.test(tmp)
res = c(resfish$estimate, unlist(resfish$conf.int), resfish$p.value)
Resdown = rbind(Resdown, res)
tmp=table(Ranges$cpg=="cpg", signif=index)
resfish=fisher.test(tmp)
res = c(resfish$estimate, unlist(resfish$conf.int), resfish$p.value)
Resdown = rbind(Resdown, res)
colnames(Resdown)=c("OR", "CI95L", "CI95U", "P")
rownames(Resdown)=c(unique(Ranges$feature), "TF-binding", "CpG-island")
Resdown$Beta = log(Resdown$OR)
Resdown$SE = (log(Resdown$OR)-log(Resdown$CI95L))/1.96
Resdown$Padj=p.adjust(Resdown$P, method = "bonferroni")

Resup=data.frame()
index = Ranges$WaldPvalue_groupCD<=thresholdp & Ranges$groupCD>thresholdLFC
for (feat in unique(Ranges$feature)){
  tmp=table(Ranges$feature == feat, signif=index)
  resfish=fisher.test(tmp)
  res = c(resfish$estimate, unlist(resfish$conf.int), resfish$p.value)
  Resup = rbind(Resup, res)
}
tmp=table(Ranges$tf_binding!="", signif=index)
resfish=fisher.test(tmp)
res = c(resfish$estimate, unlist(resfish$conf.int), resfish$p.value)
Resup = rbind(Resup, res)
tmp=table(Ranges$cpg=="cpg", signif=index)
resfish=fisher.test(tmp)
res = c(resfish$estimate, unlist(resfish$conf.int), resfish$p.value)
Resup = rbind(Resup, res)
colnames(Resup)=c("OR", "CI95L", "CI95U", "P")
rownames(Resup)=c(unique(Ranges$feature), "TF-binding", "CpG-island")
Resup$Beta = log(Resup$OR)
Resup$SE = (log(Resup$OR)-log(Resup$CI95L))/1.96
Resup$Padj=p.adjust(Resup$P, method = "bonferroni")

par(mfrow=c(1,3))

multiORplot(Resall, Pval = "P", Padj = "Padj", beta="Beta",SE = "SE", pheno="All diff. methylated loci")
multiORplot(Resup, Pval = "P", Padj = "Padj", beta="Beta",SE = "SE", pheno="hypomethylated loci")
multiORplot(Resdown, Pval = "P", Padj = "Padj", beta="Beta",SE = "SE", pheno="Hypermethylated loci")




```


## GO-term Enrichment

Significant loci and differentially methylated regions with a p-value <= `r thresholdp` and an absolute log2 fold-change lager `r thresholdLFC` were tested for enrichment among  GO-terms Molecular Function, Cellular Compartment and Biological Processes, 
KEGG pathways, Transcription factor Binding sites, Human Protein Atlas Tissue Expression, Human Phenotypes.  



```{r go_term, fig.height=15, fig.width=9}

getGOresults = function(geneset, genereference){
  resgo = gost(geneset, organism = "hsapiens",
               correction_method = "fdr",
               domain_scope = "custom",
               sources = c("GO:BP", "GO:MF", "GO:CC", "KEGG", "TF", "HP", "HPA"),
               custom_bg = genereference)
  if(length(resgo) != 0){
    return(resgo)
  } else {
    print("no significant results")
    return(NULL)
  }  
}

gene_univers = getuniquegenes(as.data.frame(rowRanges(dds_filt))$gene)

# idx = (results_Deseq$pvalue <= thresholdp) & (results_Deseq$log2FoldChange > thresholdLFC)
# genes_pos_beta = getuniquegenes(as.data.frame(rowRanges(dds_filt))$gene[idx])
# 
# idx = (results_Deseq$pvalue <= thresholdp) & (results_Deseq$log2FoldChange <= -thresholdLFC)
# genes_neg_beta = getuniquegenes(as.data.frame(rowRanges(dds_filt))$gene[idx])

idx = (results_Deseq$pvalue <= thresholdp & 
         (abs(results_Deseq$log2FoldChange) > thresholdLFC))

genes_reg = getuniquegenes(as.data.frame(rowRanges(dds_filt))$gene[idx])


dmr_genes = unique(resultsdmr_table$name[resultsdmr_table$p.value<=thresholdp & 
                   abs(resultsdmr_table$value)>=thresholdLFC])


Genes_of_interset = list("01_dmregions" = dmr_genes,  
                         "02_dmtag" = genes_reg
                         # , 
                         # "03_hypermethlated" = genes_neg_beta, 
                         # "04_hypomethlated" = genes_pos_beta
                         )

gostres = getGOresults(Genes_of_interset, gene_univers)

gostplot(gostres, capped = TRUE, interactive = T)
p = gostplot(gostres, capped = TRUE, interactive = F)

toptab = gostres$result

pp = publish_gostplot(p, filename = paste0(Home,"/output/gostres.pdf"))
write.xlsx2(toptab, file = paste0(Home,"/output/GOres.xlsx"), sheetName = "GO_enrichment")


```

## Kang Brain Module Enrichments 

Gene sets identified to be deferentially methylated with a p-value <= `r thresholdp` and an absolute log2 fold-change larger `r thresholdLFC` were tested for enrichment among gene-modules coregulated during Brain expression. 


```{r brainModules, fig.height=12, fig.width=8}

# define Reference Universe 

KangUnivers<- read.table(paste0(Home,"/data/KangUnivers.txt"), sep="\t", header=T)
colnames(KangUnivers)<-c("EntrezId","Symbol")

Kang_genes<-read.table(paste0(Home,"/data/Kang_dataset_genesMod_version2.txt"),sep="\t",header=TRUE)

#3)Generate Gene universe to be used for single gene lists
tmp=merge(KangUnivers,Kang_genes,by.y="EntrezGene",by.x="EntrezId",all=TRUE) #18826
KangUni_Final<-tmp[duplicated(tmp$EntrezId)==FALSE,] #18675


# Local analysis gene universe
Annotation_list<-data.frame(Symbol = gene_univers)

# match modules 
Annotation_list$Module = Kang_genes$Module[match(Annotation_list$Symbol,Kang_genes$symbol)]

# check if overlapping in gene universes
Annotation_list$univers = Annotation_list$Symbol %in% KangUni_Final$Symbol

# drop duplicates 
Annotation_list = Annotation_list[duplicated(Annotation_list$Symbol)==FALSE,]

# selct only genes that have been detected on both datasets
Annotation_list = Annotation_list[Annotation_list$univers==T,] 

# final reference 
UniversalGeneset=Annotation_list$Symbol

# define Gene lists to test 
# sort and order Modules to be teste 

Modules=unique(Annotation_list$Module)
Modules = Modules[! Modules %in% c(NA, "")]
Modules = Modules[order(as.numeric(gsub("M","",Modules)))]

GL_all=list()

for(i in Modules){
  GL_all[[i]]=Annotation_list$Symbol[Annotation_list$Module%in%i]
}
GL_all[["M_all"]]=Kang_genes$symbol[Kang_genes$Module %in% Modules]


GOI1 = Genes_of_interset

Resultsall=list()
for(j in names(GOI1)){
  Res = data.frame()
  for(i in names(GL_all)){
    Modulegene=GL_all[[i]]
    Factorgene=GOI1[[j]]
    Testframe<-fisher.test(table(factor(UniversalGeneset %in% Factorgene,levels=c("TRUE","FALSE")),
                                 factor(UniversalGeneset %in% Modulegene,levels=c("TRUE","FALSE"))))
    beta=log(Testframe$estimate)
    Res[i, "beta"] =beta
    Res[i, "SE"]=abs(beta-log(Testframe$conf.int[1]))/1.96
    Res[i, "Pval"]=Testframe$p.value
    Res[i, "OR"]=(Testframe$estimate)
    Res[i, "ORL"]=(Testframe$conf.int[1])
    Res[i, "ORU"]=(Testframe$conf.int[2])
  }
  Res$Padj = p.adjust(Res$Pval, method = "bonferroni")
  Resultsall[[j]] = Res
  
}
par(mfrow = c(2,2))
for (i in names(Resultsall)){ 
  multiORplot(datatoplot = Resultsall[[i]], pheno=i)
}

Modsig = c()

for(r in names(Resultsall)){
  a=rownames(Resultsall[[r]])[Resultsall[[r]]$Padj<=0.05]
  Modsig = c(Modsig,a)
}

# show brains TODO
Modisg=unique(Modsig)
load(paste0(Home,"/data/Kang_DataPreprocessing.RData")) #Load the Kang expression data of all genes 
datExprPlot=matriz #Expression data of Kang loaded as Rdata object DataPreprocessing.RData



```


## SEM

```{r SEM, , fig.width=15, fig.height=8, results = 'asis'}

dropfact=c("site")


Patdata=as.data.frame(colData(dds_filt))

load(paste0(Home, "/output/envFact.RData"))
load(paste0(Home, "/output/modelFact.RData"))

envFact=envFact[!envFact %in% dropfact] 
modelFact=modelFact[!modelFact %in% dropfact] 

EpiMarker = c()

# TopHit
Patdata$Epi_TopHit=log2_cpm[base::which.min(results_Deseq$pvalue),]

# 1PC of all diff met
tmp=glmpca(log2_cpm[base::which(results_Deseq$pvalue<=thresholdp),], 1)

Patdata$Epi_all= tmp$factors$dim1
  
EpiMarker = c(EpiMarker, "Epi_TopHit", "Epi_all")

#Brain Modules

Epitestset=GL_all[Modisg]

for(n in names(Epitestset)){
  index=gettaglistforgenelist(genelist = Epitestset[[n]], dds_filt)
  index = base::intersect(index, base::which(results_Deseq$pvalue<=thresholdp))
  # get eigenvalue
  epiname=paste0("Epi_",n)
  tmp=glmpca(log2_cpm[index,], 1)
  Patdata[,epiname]= tmp$factors$dim1
  EpiMarker = c(EpiMarker, epiname)
}

cormat = cor(apply(Patdata[,c("group", envFact, modelFact, EpiMarker)] %>% mutate_all(as.numeric), 2, minmax_scaling),
             use = "pairwise.complete.obs")

par(mfrow=c(1,2))
corrplot(cormat, main="correlations")
corrplot(cormat, order = "hclust", main="scaled correlations")

fullmodEnv=paste(envFact,modelFact, sep = "+", collapse = "+")

Dataset = Patdata[,c("group", envFact,modelFact,EpiMarker)]

model = paste0("group ~ Epi+", fullmodEnv,"
Epi ~" , fullmodEnv,"
Epi~~Epi
group~~group
")

Netlist = list()

for (marker in EpiMarker) {
  Dataset$Epi = Dataset[,marker]
  Datasetscaled = Dataset %>% mutate_if(is.numeric, minmax_scaling)
  Datasetscaled = Datasetscaled %>% mutate_if(is.factor, ~ as.numeric(.)-1)
  
  fit<-lavaan(model,data=Datasetscaled)
  
  sink(paste0(Home,"/output/SEM_summary_group",marker,".txt"))
  summary(fit)
  fitMeasures(fit)
  sink()
  
  #SOURCE FOR PLOT https://stackoverflow.com/questions/51270032/how-can-i-display-only-significant-path-lines-on-a-path-diagram-r-lavaan-sem
  restab=lavaan::standardizedSolution(fit) %>% dplyr::filter(!is.na(pvalue)) %>% 
    arrange(desc(pvalue)) %>% mutate_if("is.numeric","round",3) %>% 
    dplyr::select(-ci.lower,-ci.upper,-z)
  
  pvalue_cutoff <- 0.05
  
  obj <- semPlot:::semPlotModel(fit)
  original_Pars <- obj@Pars
  check_Pars <- obj@Pars %>% dplyr:::filter(!(edge %in% c("int","<->") | lhs == rhs)) # this is the list of parameter to sift thru
  keep_Pars <- obj@Pars %>% dplyr:::filter(edge %in% c("int","<->") | lhs == rhs) # this is the list of parameter to keep asis
  
  test_against <- lavaan::standardizedSolution(fit) %>% dplyr::filter(pvalue < pvalue_cutoff, rhs != lhs)
  
  # for some reason, the rhs and lhs are reversed in the standardizedSolution() output, for some of the values
  # I'll have to reverse it myself, and test against both orders
  
  test_against_rev <- test_against %>% dplyr::rename(rhs2 = lhs, lhs = rhs) %>% dplyr::rename(rhs = rhs2)
  
  checked_Pars <-
    check_Pars %>% semi_join(test_against, by = c("lhs", "rhs")) %>% bind_rows(
      check_Pars %>% semi_join(test_against_rev, by = c("lhs", "rhs"))
    )
  
  obj@Pars <- keep_Pars %>% bind_rows(checked_Pars) %>% 
    mutate_if("is.numeric","round",3) %>% 
    mutate_at(c("lhs","rhs"),~gsub("Epi", marker,.))
  
  obj@Vars = obj@Vars %>% mutate_at(c("name"),~gsub("Epi", marker,.))
  
  DF = obj@Pars
  DF = DF[DF$lhs!=DF$rhs,]
  DF = DF[abs(DF$std)>0.1,]
  
  DF = DF[DF$edge == "~>",] # only include directly modelled effects in figure 
  
  nodes <- data.frame(id=obj@Vars$name, label = obj@Vars$name)
  nodes$color<-Dark8[8]
  nodes$color[nodes$label == "group"] = Dark8[3]
  nodes$color[nodes$label == marker] = Dark8[4]
  nodes$color[nodes$label %in% envFact] = Dark8[5]
  
  edges <- data.frame(from = DF$lhs, 
                      to = DF$rhs, 
                      width=abs(DF$std), 
                      arrows ="to")
  
  edges$dashes = F
  edges$label =  DF$std
  edges$color=Dark8[1:2][factor(sign(DF$std), levels=c(-1,0,1),labels=c(1,2,2))]
  plotnet<- visNetwork(nodes, edges,
                       main=list(text=marker,
                                 style="font-family:arial;font-size:20px;text-align:center"),
                       submain=list(text="significant paths", style="font-family:arial;text-align:center")) %>% 
    visEdges(arrows =list(to = list(enabled = TRUE, scaleFactor = 0.7))) %>%
    visPhysics(enabled = T, solver = "forceAtlas2Based")
  
  Netlist[[marker]] = plotnet
  htmlfile = paste0(Home,"/output/SEMplot_",marker,".html")
  visSave(plotnet, htmlfile)
  webshot(paste0(Home,"/output/SEMplot_",marker,".html"), zoom = 1, 
          file = paste0(Home,"/output/SEMplot_",marker,".png"))
}

```

```{r plotsems_prep, results="asis"}

rmd_paths <-paste0(tempfile(c(names(Netlist))),".Rmd")
names(rmd_paths) <- names(Netlist)

for (n in names(rmd_paths)) {
    sink(file = rmd_paths[n])
    cat("  \n",
        "```{r, echo = FALSE}",
            "Netlist[[n]]",
        "```",
        sep = "  \n")
    sink()
}

```

## Interactive results SEM analysis 

only direct effects with a significant standardized effect of p<0.05 are shown. 


```{r plotsems_show_interactive, results='asis'}
    for (n in names(rmd_paths)) {
        cat(knitr::knit_child(rmd_paths[[n]],
                              quiet= TRUE))
        file.remove(rmd_paths[[n]])
    }
```





